\thetitle

### 2020-11-19 writing/introduction.tex
small edits

The paper's second contribution is a set of results from applying the evaluation
method to three distinct blame assignment strategies and approximately
72000 cases (section~\ref{sec:results}): Vitousek et
al.'s strategy of tracking typed/untyped boundaries in Reticulated;
Tobin-Hochstadt et al.'s use of higher-order contract system~\cite{ff-icfp-2002,
mf-toplas-2009}; and the approach of industrial systems that forego blame
and instead rely on error messages from the
underlying programming language.

\mf{This is made up and must be fixed for the submission.}

A direct interpretation of the results (section~\ref{sec:discussion}) is that
they validate the conjecture behind the work of theoreticians. A good blame assignment
strategy
greatly shortens the search for impedance mismatches between the specified types
and the behavior of untyped components. Even better, the evaluation method clearly
identifies the wrapper-based blame tracking of Typed Racket as the vastly superior
academic system. Vitousek et al. lose one more time. Of course, a detailed
discussion must acknowledge the threats to validity of these results.

>>>>

The paper's second contribution is a set of results from applying the evaluation
method to three distinct blame assignment strategies and approximately
72000 cases (section~\ref{sec:results}): Vitousek et
al.'s strategy of tracking typed/untyped boundaries in Reticulated;
Tobin-Hochstadt et al.'s use of higher-order contract system~\cite{ff-icfp-2002,
mf-toplas-2009}; and the approach of industrial systems that forego blame
and instead rely on error messages from the
underlying programming language. A direct interpretation of the results
(section~\ref{sec:discussion}) is that
they validate the conjecture behind the work of theoreticians. A good blame assignment
strategy helps the search for impedance mismatches between the specified types
and the behavior of untyped components. Even better, the evaluation method 
identifies the wrapper-based blame tracking of Typed Racket as the most
useful academic system with that of Reticulated Python just a short
distance behind.

### 2020-11-19 writing/paper.tex
ogranizing section 5
format + prose tweaks
revision of rational
small edits

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,screen,anonymous,nonacm]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
%% For double-blind review submission, w/ CCS and ACM Reference
%%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

@@

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
%% \setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%
\setcopyright{none}
\acmPrice{1}
\acmDOI{}
\acmYear{}
\copyrightyear{}
\acmJournal{pacmpl}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmMonth{1}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Programming language theoreticians develop blame assignment systems
and prove blame theorems for gradually typed programming languages.
Practical implementations of gradual typing almost completely ignore
the idea of blame assignment.  This contrast raises the question
whether blame assignment provides any value to the working programmer
and poses the challenge of how to evaluate the effectiveness of blame
assignment strategies. This paper contributes (1) the first evaluation
method for blame assignment strategies and (2) the results from applying
it to three different gradual type systems. \mf{This sentence must be checked
before final submission.} These results confirm that blame assignment
is helpful to a rationally acting programmer.
\end{abstract}

@@

\input{introduction.tex} 
\input{challenges.tex}   
\input{landscape.tex} 
\input{mutate.tex}
\input{rational.tex}
\input{sample.tex}
\input{results.tex}      \label{sec:results}
\input{discussion.tex}   \label{sec:discussion}
\input{related.tex}      \label{sec:related}
\input{conclusion.tex}   \label{sec:conclusion}

>>>>

\documentclass[sigplan,10pt,review,anonymous]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

@@

\acmConference[]{}{}{}
\acmYear{}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Programming language theoreticians develop blame assignment systems
and prove blame theorems for gradually typed programming languages.
Practical implementations of gradual typing almost completely ignore
the idea of blame assignment.  This contrast raises the question
whether blame assignment provides any value to the working programmer
and poses the challenge of how to evaluate the effectiveness of blame
assignment strategies. This paper contributes (1) the first evaluation
method for blame assignment strategies and (2) the results from applying
it to three different gradual type systems. These results confirm that blame assignment
is helpful to a rationally acting programmer.
\end{abstract}

@@

\input{introduction.tex}    \label{sec:introduction}
\input{challenges.tex}      \label{sec:challenges}
\input{landscape.tex}       \label{sec:landscape}
\input{mutate.tex}          \label{sec:mutate}
\input{rational.tex}	      \label{sec:rational}
\input{sample.tex}          \label{sec:sample}
\input{results.tex}         \label{sec:results}
\input{discussion.tex}      \label{sec:discussion}
\input{related.tex}         \label{sec:related}
\input{conclusion.tex}      \label{sec:conclusion}

### 2020-11-19 writing/sample.tex
small edits

As outlined in section~\ref{sec:rational}, all our experimental questions
require that we decide whether a mode of the rational programmer is
always, sometimes or never useful for all the interesting scenarios of a
mutant $\system$. Assuming a normal distribution for the 
interesting scenarios in the lattice $\lattice{\system}$, we can obtain an
estimate answer with high confidence by randomly picking a sufficient
number of interesting scenarios $\lattice{\system}$. Thus
$\lattice{\system}$ becomes the first stratum for our sampling. For
95\% confidence, that is a 5\% margin, the sample size needs
to be just 95 out of $2^N$ scenarios in $\lattice{\system}$ where $N$ is the
number of components of $\system$.

The second sampling is the set of mutants. After all, our mutators
generate 16,800 mutants with
interesting scenarios; still far too many to explore even after sampling
the lattice of each mutant.  Unlike the lattice of scenarios of a mutant
, the set of mutants is not a coherent and finite population;  it is 
unclear how we can relate a sample of mutants to the
of all possible mutants. Thus we sacrifice the ability to generalize 
beyond the mutants we randomly select and opt for the largest random
sample that keeps the experiment computationally feasible. 
In particular,  we sample 80 mutants with interesting scenarios
per benchmark and for any benchmark, we ensure that our sample reflects the diversity of mutants with respect
to the mutators that generated them. Specifically, we select the 80 mutants evenly-distributed
across all of the mutators that contribute mutants for a benchmark.

>>>>

As outlined in section~\ref{sec:rational}, all our experimental questions
require that we answer first whether a mode of the rational programmer is
useful or not for all the interesting scenarios of a mutant $\system$.
Given the binary nature of this basic question, we can rely on random
sampling of interesting scenarios from $\lattice{\system}$ to obtain an
approximate answer.  Thus $\lattice{\system}$ becomes the first stratum
for our sampling and, given a sufficiently large sample, we can get with
high confidence either a positive answer about the usefulness of a mode of
the rational programmer or a negative answer or neither, in which case we
deduce that the mode we examine is some times useful and some times it is
not. For 95\% confidence interval and 5\% error margin, the sample  needs
to include just 96 out of the $2^N$ scenarios of $\lattice{\system}$
where $N$ is the number of components of $\system$. For lattices that have
less than 96 interesting scenarios we select all their interesting
scenarios.

The second sampling stratum is the set of mutants. After all, our mutators
generate 16,800 mutants with interesting scenarios; still far too many to
explore even after sampling the lattice of each mutant. Thus we randomly
select form the mutants of each benchmarks to obtain  the largest random
sample that keeps the experiment computationally feasible  while ensuring
that, per benchmark, the sample reflects the diversity of mutants per with
respect to the mutators that generated them.  Specifically,  we sample 80
mutants with interesting scenarios per benchmark  evenly-distributed
across all of the mutators that contribute mutants for a benchmark.
